The code for this project is developed in the C programming language, utilizing the Visual Studio integrated development environment (IDE) to facilitate efficient code development and debugging.
\\To address the complexity of the optimization problem at hand, the project relies on two crucial libraries, namely CPLEX \cite{CPLEX} and Concorde. \\CPLEX is a powerful optimization library that provides various algorithms for solving linear programming, mixed-integer programming, and quadratic programming problems. It has been instrumental in formulating and solving the Traveling Salesman Problem (TSP) in this work.
\\
Concorde is a powerful library for solving the Traveling Salesman Problem (TSP) and related combinatorial optimization problems. It provides a collection of high-quality heuristics and exact algorithms to find near-optimal solutions for large-scale TSP instances. 
\\
For obtaining the input dataset for the TSP instances, we utilized some instances of the TSPLIB, which is a comprehensive library providing a range of benchmark instances for the Traveling Salesman Problem. The use of this dataset ensures standardization and comparability in our research.
\\
For visualizing the results, we employed gnuplot, a versatile and widely used plotting program. Gnuplot enabled us to generate clear and comprehensive visual representations of our solution outputs, aiding in better comprehension and analysis.
\\
In order to assess the performance and effectiveness of our implemented algorithms, we employed a performance profile tool developed by Professor Salvagnin.
