By exploring various Heuristic and Metaheuristic Algorithms, Exact methods using CPLEX models, callbacks, and the Concorde library, Matheuristic Algorithms with different kind of contraints, we have provided a comprehensive and detailed overview of different solution strategies for the complex Travelling Salesman Problem.

From the results of your comparative analysis, it is evident that each approach has its strengths and limitations. Exact algorithms, such as those implemented with CPLEX and callbacks, guarantee optimal solutions but may require significant computation time for large instances.

On the other hand, heuristic and metaheuristic algorithms have shown to be faster in obtaining acceptable solutions within reasonable time frames, even if they might not always be optimal. However, they have proven their utility in situations where the primary goal is to achieve a satisfactory solution within acceptable time constraints, such as real life problems.

The Concorde library has distinguished itself for its outstanding performance in tackling the TSP. The usage of this library with callbacks has proven to be an effective approach in further improving the solutions obtained from exact algorithms, even if it could be much more time-consuming over larger instances.

Based on the results and analyses, it can be concluded that there is no one-size-fits-all approach. The choice of the resolution method depends on the specific requirements of the problem, such as instance size, required accuracy, and available computational resources.

Additionally, it is essential to emphasize the importance of considering further research into new approaches or the combination of multiple strategies to address the TSP even more efficiently and accurately.